\section{Response to Associate Editor}\label{sec:ae}

\comment{We invite you to revise your paper, carefully addressing the comments from the reviewers and the editor.}

\answer{Thank you for the opportunity. We have thoroughly addressed all the points raised by the reviewers, as discussed in detail in the following of this reply letter. In summary: 

\begin{enumerate}
    \item we did a complete English proofread addressing all typos and grammar errors;
    \item we increased the quality and readability of all figures;
    \item we revised the abstract and clarified all assumptions in the paper;
    \item according to the following Reviewer~4's comment:
\vspace{0.5em}
       
\begin{adjustwidth}{10pt}{10pt}

''\emph{The evaluation for the proposed approach is still not enough. The performance of the proposed approach should be extensively validated. Although, the work has tested its performance on some metrics, the proposed system should be fully tested on various evaluation metrics in various scenarios so that the real performance can be verified. Considering that the real-world application is more complex, the readers could wonder if the proposed approach is practical to be applied to real-world applications (or how practical the proposed method can be applied to real-world applications). I understand that I may introduce some extra work. In the future work, the authors may consider it.}"
\end{adjustwidth}

\vspace{0.5em}

we refined and clarified the discussion on the findings and results retrieved by our experimental evaluation; we also clarified in the text our future work, with particular reference to the experimental evaluation.
\end{enumerate}
~}