\documentclass{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{csquotes}
%\usepackage{soul, color}
\usepackage[shortcuts]{extdash}
\usepackage{color}
\usepackage{graphics}
\usepackage[normalem]{ulem}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{url}
\usepackage{float}
\usepackage{xspace}
\usepackage{tabularx}


\usepackage{adjustbox}
%\usepackage[dvipsnames,table]{xcolor}
%\colorlet{OurColor}{MidnightBlue}
\usepackage[]{xcolor}
\colorlet{OurColor}{blue}
%\colorlet{Nicola}{green!70!black}
%\colorlet{MinorColor}{violet}
\usepackage[skip=2pt]{parskip}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{makecell}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{pifont}
\usepackage{url}
\usepackage{calc}
\usepackage{booktabs}
\usepackage{ifthen}

\usepackage{tikz}
\usetikzlibrary{arrows.meta, backgrounds, calc, fit, %graphs, graphdrawing,
shapes.misc}

\usepackage{soul, color}

\usepackage{geometry}
 \geometry{
 a4paper,
 %total={170mm,257mm},
 left=30mm,
 right=30mm,
 top=23mm,
}

\usepackage[hidelinks,
  pdfauthor={Antongiacomo Polimeno, Chiara Braghin, Marco Anisetti, Claudio A. Ardagna},
  pdftitle={Revision to the Paper Balancing Data Quality and Protection in Distributed Big Data Pipelines (Original Title)\\
(Current Title)},]{hyperref}

\usepackage[inline]{enumitem}
\usepackage{changepage}
\usepackage{cleveref}

\setlist*[enumerate]{label=\emph{\roman*)}}

\parindent 0in

\newcommand{\old}[1]{\sout{#1}}
\newcommand{\new}[1]{\uline{#1\/}}

\newcommand{\comment}[1]{\noindent \textbf{Comment:} #1\\}
\newcommand{\answer}[1]{\noindent \textbf{Response:} {\it #1\/}\\\hrule ~\\}
\newcommand{\answerwithout}[1]{\noindent \textbf{Response:} {\it #1\/}\\}

\newcommand{\reference}[1]{[#1]}


\begin{document}

\input{00_macro.tex}

\newcommand{\customTheoremBlock}[2]{\begin{adjustwidth}{25pt}{25pt}\textbf{\emph{#1}} #2\end{adjustwidth}}
\newcommand{\customFigCaption}[2]{{\footnotesize Fig. #1.~#2}}
\newcommand{\customSecName}[2]{{\large \textbf{#1~#2}}}
\newcommand{\customTableCaption}[2]{{\centering TABLE~#1 #2}}

\title{\large \textbf{Revision to the Paper ID b5d6583f-1abe-4325-9459-588bcec2cdad\\Maximizing Data Quality While Ensuring Data
Protection in Service-Based Data Pipelines\\}}

\author{Antongiacomo~Polimeno, Chiara~Braghin, Marco~Anisetti, Claudio~A.~Ardagna}

\date{}

\maketitle

We would like to thank the associate editor and reviewers for their valuable and helpful reviews. We revised the paper according to their comments.

\vspace{1em}

This letter is organized in three sections, one for the associate editor and each of the two  reviewers, where we provide our detailed responses and explain the rationale behind the modifications made to the manuscript.
Please note that in the reviewers' comments, sections and page numbers refer to the old version
of the paper, whereas in our replies they refer to the revised version of the paper. For reviewers' convenience, major modifications have been highlighted in {\color{OurColor}blue} in the revised paper and in this reply letter.

\vspace{1em}

As a general remark, following the reviewers' comments, we did a complete English proofread addressing all typos and grammar errors, and increasing the quality and readability of all figures. We also revised the abstract and clarified all assumptions done in the paper. We finally clarified the relevance of our experimental evaluation and the corresponding evaluation parameters, discussing our future plans to a more extensive experimental evaluation.

\newpage

\vspace{.4cm}
%----------------------------------------------------%

\input{01_ae}

\newpage
\vspace{.4cm}

%----------------------------------------------------%
%----------------------------------------------------%

\input{01_r3}

\newpage
\vspace{.4cm}

%----------------------------------------------------%

\input{02_r4}

\newpage
\vspace{.4cm}

%----------------------------------------------------%

\end{document}